---
title: "Assignment 6"
author: "Charles Lang"
date: "12/1/2020"
output: html_document
---

```{r}
library(tidyverse)
library(GGally)

```

Instructions

Task I
You will be working in groups on a quantified student project
Download the Sensor Kinetics Pro app to your iOS or Android device (or similar gyroscope measuring app)
Split into subgroups with each subgroup investigating one measure available in the app




Our awesome activity is as follows...
  Go up the stairs, [speed]
  while shouting, [audio]
  while pressing on the screen [pressure]
  while tossing and catching a ball 5 times 
  What counts as ggood is doing it the fastest (least time)

App: phyphox on the App Store https://apps.apple.com/us/app/phyphox/id1127319693


Measure the following sensors:
  Pressure - pressing on screen
  Audio - audio scope
  Elevator - speed


Avg speed:
  Berj Akian ba2629   ~5 seconds 
  Fei Wang fw2347		  ~7 seconds 
  Xiaojia Liu xl2958	~9 seconds
  Guoliang Xu gx2150	~10 seconds
  Jiacong Zhu jz3254	~10	seconds
  Xijia Wang xw2653		~10 seconds 
  Yingxin Xie yx2614 	~12 seconds



Each subgroup should report back after 20min with what they have learned




Task II
In consultation with your group select a physical activity that: 
a) can be measured using the app, 

b) has a concrete, continuous, countable outcome and 

c) some members of the group are experts and some are novices at - If you are looking for inspiration you can find previous student projects below.

Collect measurements for all members of the group performing the task using the app

Create a Github repo where you can store and share your data
```{r}
#---- read in the data
DF1 <- read.csv("stairsalldata1.csv")

#---- what do we have?...
str(DF1)
head(DF1)
#summary(DF1)

#---- na's please rise... https://sebastiansauer.github.io/sum-isna/
DF1.is.na <- DF1 %>% 
  summarise_all(funs(sum(is.na(.))))
DF1.is.na

sum(DF1.is.na)

#---- drop rows that contain NA in any column... if needed based on results of previous command... not needed
#DF2 <- na.omit(D1)

```



Using K-means, can you identify the novices from the experts using only the app data?
```{r}
#--- Clustering, for fun

#--- what class is the data
class(DF1)


#--- a little more cleanup
DF2 <- DF1[,3:5]
rownames(DF2) <- DF1[,1]

 
#--- convert to matrix, if needed
DF2.matrix <- as.matrix(DF2)

 
#--- how many unique rows... this is the upper limit to k values
dim(unique(DF2.matrix))


#--- calculate Total within-cluster sum of squares 
DF2.tot_withinss <- map_dbl(1:5,  function(k){
  model <- kmeans(DF2.matrix, centers = k)
  model$tot.withinss
})

 
#--- Generate a data frame containing both k and tot_withinss
DF2.elbow_df <- data.frame(
  k = 1:5,
  tot_withinss = DF2.tot_withinss
)


#--- Plot the elbow plot of K values and tot_withinss
ggplot(DF2.elbow_df, aes(x = k, y = tot_withinss)) +
  geom_line() +
  scale_x_continuous(breaks = 1:5)
 

#--- the elbow plot looks like 2 groups... maybe 3 groups for good measure
 

#--- kmeans the datatable into 2 groups 
DF2.groups <- kmeans(DF2.matrix, centers = 2)
 
DF2.clustered <- data.frame(DF2, cluster = as.factor(DF2.groups$cluster))


DF2.clustered <- cbind(DF2.clustered,time = DF1[,2])


#--- Can you identify the novices from the experts using only the app data? 
#--- pressure: generally, higher pressure value was associated with slowest speed
#--- audio: generally, lower audio value was associated with slower speed

```


Visualize your results using ggplot2

```{r}
#--- Let's take a look

DF3.clustered <- DF2.clustered
DF3.clustered$cluster <- as.numeric(DF3.clustered$cluster)

pairs(DF3.clustered)
  
ggpairs(DF3.clustered,
          progress = FALSE,
          lower = list(continuous = wrap("points", alpha = 0.3, size=0.2)))
  
ggpairs(DF3.clustered,
          progress = FALSE,
          lower = list(continuous = wrap("smooth", alpha = 0.3, size=0.2)))

```



Task III
Devise five questions on a five point Likert scale that ask members of your group about their experience with your chosen task
```{r}

#--- Question 1: the task was... fun
#--- Question 2: the task was... easy
#--- Question 3: the task was... confusing
#--- Question 4: the task was... fast
#--- Question 5: the task was... embarrassing

#--- (1) Strongly disagree; (2) Disagree; (3) Neither agree nor disagree; (4) Agree; (5) Strongly agree
```



Collect data from all members of your group
```{r}
#---- read in the data
SV1 <- read.csv("survey1.csv")

#---- what do we have?...
str(SV1)
head(SV1)
#summary(DF1)

#---- na's please rise... https://sebastiansauer.github.io/sum-isna/
SV1.is.na <- DF1 %>% 
  summarise_all(funs(sum(is.na(.))))
SV1.is.na

sum(SV1.is.na)

#---- drop rows that contain NA in any column... if needed based on results of previous command... not needed
#SV2 <- na.omit(SV1)
```

Run a PCA analysis on the data
```{r}

#--- commence dimension reduction
SV1.pca <- prcomp(SV1[2:6], scale. = TRUE, center = TRUE)

SV1.pca$sdev

SV1.pca$sdev^2

summary(SV1.pca)

plot(SV1.pca, type = "lines")

biplot(SV1.pca)

ggbiplot(SV1.pca)

#--- Now print out the loadings for the components you generated:

SV1.loadings <- abs(SV1.pca$rotation)

SV1.loadings


```

What does PC1 represent in your analysis?
```{r}

#PC1 represents the greatest diversity of the survey response data.... 0 points

#PC1 represents the following proportionate combination of survey responses (absolute value of the rotation values >> loadings)...0 points
#Q4 0.5797535
#Q2 0.5755824
#Q1 0.5517591
#Q5 0.1261263
#Q3 0.1106564

#PC1 represents ~47% of the total variance in the data... 0 points



#PC1 has a disproportionate representation of questions 4, 2, 1 



#The similarity in the proportional representation of questions 4, 2, 1 represents a commonality of the theme of those questions. 

#--- it would seem these three questions have a common theme of 'joy', 'positive', or 'satisfying' theme... thus they would be answered similarly
#--- Question 4: the task was... fast
#--- Question 2: the task was... easy
#--- Question 1: the task was...fun



#The similarity in the proportional representation of questions 5 and 3 represents a commonality of the theme of those questions as well. 

#--- it would seem these three questions have a theme of 'not-joyful', 'negative', or 'gloomy' theme... thus they would be answered similarly
#--- Question 5: the task was... embarrassing
#--- Question 3: the task was... confusing


#--- so neat that PC analysis found those thematic relationships between the questions and the answers!!!!




```

Visualize the PC1 results for each member of your group
```{r}
#--- create a dataframe that brings together person name, PC1, and PC2... (and round those pesky long PC values to 4 places after the decimal)
SV2 <- as.data.frame(cbind(person = SV1$Person, 
                           PC1 = round(SV1.pca$x[,1],4), 
                           PC2 = round(SV1.pca$x[,2],4)
                           )
                     )

#--- Thee Oh Sees - Ugly Man ♫...
plot(SV2[,2])


#--- Can we make a better graph? Yes we can! 

#--- turns out the PC values are stored as character strings... and so they don't sort numerically... we can fix this by sorting the whole dataframe as if PC1 was a number... and then adding a column that has the sort order... we'll use that later, muahahaha
SV2 <- SV2[order(as.numeric(SV2$PC1), decreasing = TRUE),]
SV2$sort <- 1:7


#--- Mona Lisa... Many dreams have been brought to your doorstep ♫...
x <- reorder(SV2$person,SV2$sort)
y <- as.numeric(SV2$PC1)

ggplot(SV2, aes(x, y, label = x, color = x)) +
        geom_point()  +
        geom_text(aes(label = y),vjust = -.5, size = 3, fontface = 4) +
        theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
        ggtitle("PC1 for all survey respondents") +
        xlab("Survey Respondents") +
        ylab("PC1") +
        theme(legend.position = "none") + 
        geom_hline(yintercept = 0, linetype = "dotted")



```



Task IV
If you could only choose one thing to predict the score for each member of your group (raw data variable, a cluster or a PC) what would it be and why?
```{r}


#---correlate the pc values to the activity 
#---bring together... SV1.pca$x[SV2]...  cluster membership[DF2.clustered]... 5 survey measures[SV1], time and sensor measures[DF1]

E1 <- as.data.frame(cbind(person = SV1$Person,
                          PC1 = round(SV1.pca$x[,1],4), 
                          cluster = as.numeric(DF2.clustered$cluster),
                          SV1[,2:6],
                          DF1[,2:5])
                     )

#---add a sort of PC1 
E2 <- E1[order(as.numeric(E1$PC1), decreasing = TRUE),]
E2$PC1sort <- 1:7



#---view a ggpairs correlation 
ggpairs(E2[,c(2,3,9)], progress = FALSE)
ggpairs(E2[,4:12], progress = FALSE)




#---choose one thing to predict the score for each member of your group....
#---the variable with the greatest positive correlation with greater Time (our goal is less time) is PC1... a positive correlation of 0.703
#---the variable with the greatest negative correlation with a greater Time (our goal is less time) is Question4 "the task was... fast"... a negative correlation of -0.764





```

Create a visualization that supports your conclusion
```{r}

#---see ggpairs plots above


```

Based on your conclusion devise an intervention that would help novice members of your group improve their performance
```{r}


#---since neither question 4 or PC1 can be directly controlled to change the performance...
#---the next best thing is to look at the sensor data... 
#---pressure was a negative correlation of -.442 with greater time (our goal is less time)
#---an intervention might be to tell participants to 'not push so hard on the sensor while trying to move more quickly up the stairs' 


```



Task V
Submit a link to your Github repo containing your data, code, visualizations and conclusions to charles.lang@tc.columbia.edu by 5pm EDT 12/14/20.


```{r}

#--- this assignment was ggreat... learned a lot!!
#--- here's our shared doc with our plan  
#--- https://docs.google.com/document/d/1jyhde406-YOxZl_-Smd3LRbg1BhcxnSQfl4cUJU65iQ/edit?usp=sharing

```


ggood luck!

